---
title: "STAT/MATH 495: Problem Set 07"
author: "Luke Haggerty"
date: "2017-10-24"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
    collapsed: false
    smooth_scroll: false
    df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, fig.width=8, fig.height=4.5, message=FALSE, warning = FALSE
  )
set.seed(76)

# Load packages
library(tidyverse)
library(broom)
library(knitr)
library(ROCR)

train <- read_csv("data/cs-training.csv") %>% 
  rename(Id = X1)
test <- read_csv("data/cs-test.csv") %>% 
  rename(Id = X1)
submission <- read_csv("data/sampleEntry.csv")
```

Information on the competition can be found [here](https://www.kaggle.com/c/GiveMeSomeCredit/data).



# Collaboration

Please indicate who you collaborated with on this assignment: I worked individually.



# Build binary classifier

Build the binary classifier based on a single predictor variable: `DebtRatio`,
`age`, or `MonthlyIncome`. Justify this choice.

```{r}
ggplot( data = train, aes(x = age)) + geom_histogram(binwidth = 4.4)
summary(train$age)

ggplot( data = train, aes(x = DebtRatio)) + geom_histogram(binwidth = 3300)
summary(train$DebtRatio)

ggplot( data = train, aes(x = MonthlyIncome)) + geom_histogram(binwidth = 30000)
summary(train$MonthlyIncome)

```

I examined the three variables with histograms and five number summaries. Both MonthlyIncome and DebtRatio are highly skewed. I decided to use age because it is more normally distributed.


```{r}
model_logistic <- glm(SeriousDlqin2yrs ~age,family=binomial, data=train)

log_odds_hat <- predict(model_logistic, newdata=test)
p_hat <- 1/(1 + exp(-log_odds_hat))

fitted_model <- model_logistic %>% 
  broom::augment() %>% 
  as_tibble() %>% 
  mutate(p_hat = 1/(1 + exp(-.fitted)))
predictions <- model_logistic %>% 
  broom::augment(newdata=test) %>% 
  mutate(p_hat = 1/(1 + exp(-.fitted)))

ggplot(NULL) +
  geom_line(data=fitted_model, aes(x=age, y=.fitted), col="blue") +
  geom_point(data=predictions, aes(x=age, y=.fitted), col="red")
```

# ROC curve

Based on the ultimate classifier you choose, plot a corresponding ROC curve.

```{r}
train_augmented <- model_logistic %>% 
  broom::augment() %>% 
  as_tibble() %>% 
  mutate(p_hat = 1/(1+exp(-.fitted)))

pred <- prediction(predictions = train_augmented$p_hat, labels = train_augmented$SeriousDlqin2yrs)
perf <- performance(pred, "tpr","fpr")

auc <- as.numeric(performance(pred,"auc")@y.values)
auc

plot(perf, main=paste("Area Under the Curve =", round(auc, 3)))
abline(c(0, 1), lty=2)
```

```{r}
submission$Probability <- p_hat
write.csv(submission, "submission.csv", row.names = FALSE)
```


# ROC curve for random guessing

Instead of using any predictor information as you did above, switch your
predictions to random guesses and plot the resulting ROC curve.

```{r}
rand<-runif(150000,0,1)

predr <- prediction(rand, labels = train_augmented$SeriousDlqin2yrs)
perfr <- performance(pred, "tpr","fpr")

plot(perfr, main=paste("Area Under the Curve =", round(auc, 3)))
abline(c(0, 1), lty=2)
```
